Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-17 paper.tex
Update paper.tex

\usepackage[normalem]{ulem}

@@

@@

To take advantage of multiple-core architecture of modern CPUs, an operating 
system (OS) is often designed to be highly concurrent, and processes shared 
data structures in parallel. To ensure safe access to shared data structures, 
fine-grained locking mechanisms are used to avoid data races among different 
threads. However, it is hard to determine whether a lock should be used for a 
specific data structure field (locking rules) even for an expert developer, due 
to poor documentation and complicated logic of OS code. As a result, OS kernel 
is prone to data races, because necessary locks may be missed by mistake. 
Static analysis is a common technique to help improve code quality, but it is 
quite challenging to detect data races in OS kernels automatically, because of 
lack of knowledge of locking rules and high complexity of concurrent execution.

In this paper, we design a practical static analysis approach named RaceMiner, 
to effectively detect harmful races that can trigger memory or logical bugs in 
OS kernels by mining locking rules. RaceMiner first employs an alias-aware rule 
mining method to automatically deduce locking rules, and detects data races 
caused by violation of these rules. And then performs a lock-usage analysis to 
filter out false positives caused by concurrency. At last, RaceMiner extracts 
harmful data races from all detected data races through a pattern-based 
estimation. We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 
with a false positive rate of xxx. Among these data races, xxx are estimated to 
be harmful. We have reported these harmful bugs to Linux kernel developers, and 
xxx of them have been confirmed.

>>>>

\usepackage[normalem]{ulem}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{xspace}

@@

\newcommand{\sys}{\mbox{RaceMiner}\xspace}

@@

To take advantage of multiple-core architecture of modern CPUs, an operating 
system (OS) is expected to handle concurrency, inevitably introducing many 
concurrent issues. And data race is one of the most harmful concurrent issues 
that can trigger memory or logical bugs. A data race is caused by wrongly 
accessing shared data in different threads. To ensure correct access to shared 
data, fine-grained locking mechanisms are introduced. However, it is hard to 
determine whether a lock should be used for a specific data structure field 
(locking rules) even for an expert developer, due to poor documentation and 
complicated logic of OS code. As a result, OS kernel is prone to data races, 
because necessary locks may be missed by mistake. Static analysis is a common 
technique to help improve code quality, but it is quite challenging to detect 
data races in OS kernels automatically, because of lack of knowledge of locking 
rules and high complexity of concurrent execution.

In this paper, we design a practical static analysis approach named \sys, to 
effectively detect harmful races that can trigger memory or logical bugs in OS 
kernels by mining locking rules. \sys first employs an alias-aware rule mining 
method to automatically deduce locking rules, and detects data races caused by 
violation of these rules. And then performs a lock-usage analysis to filter out 
false positives caused by concurrency. At last, \sys extracts harmful data 
races from all detected data races through a pattern-based estimation. We have 
evaluated \sys on Linux 6.2, and find xxx data races, with a false positive 
rate of xxx. Among these data races, xxx are estimated to be harmful. We have 
reported these harmful bugs to Linux kernel developers, and xxx of them have 
been confirmed.
